---
title: "Predictions using the given Dataset"
author: "Raghavandaar"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lattice)
library(ggplot2)
library(plyr)
library(randomForest)
library(caret)
library(rpart)
library(rpart.plot)
library(corrplot)

```

## Executive Summary

We will try to perform partical machine learning on the dataset



We will be doing the following:

 Process the data
 Explore the data
 Data Modeling and predicting

## Processing

We are going to process the data from the dataset
```{r}
trainRaw <- read.csv("pml-training.csv")
testRaw <- read.csv("pml-testing.csv")
```

## Exploratory data analyses 

We will try to analyse and explore the data
```{r}
dim(trainRaw)
dim(testRaw)
```
A lot of NA values present. Therefore we remove that.
```{r, cache = T}
sum(complete.cases(trainRaw))
```
```{r, cache = T}
trainRaw <- trainRaw[, colSums(is.na(trainRaw)) == 0] 
testRaw <- testRaw[, colSums(is.na(testRaw)) == 0] 
```  
```{r, cache = T}
classe <- trainRaw$classe
trainRemove <- grepl("^X|timestamp|window", names(trainRaw))
trainRaw <- trainRaw[, !trainRemove]
trainCleaned <- trainRaw[, sapply(trainRaw, is.numeric)]
trainCleaned$classe <- classe
testRemove <- grepl("^X|timestamp|window", names(testRaw))
testRaw <- testRaw[, !testRemove]
testCleaned <- testRaw[, sapply(testRaw, is.numeric)]
```

```{r, cache = T}
set.seed(22519) 
inTrain <- createDataPartition(trainCleaned$classe, p=0.70, list=F)
trainData <- trainCleaned[inTrain, ]
testData <- trainCleaned[-inTrain, ]
```

The data has been explored and cleansed. We now move on to model selection.

## Data Modeling and predicting 

We will try to modify the dataset for model selection using random forest

```{r, cache = T}
controlRf <- trainControl(method="cv", 5)
modelRf <- train(classe ~ ., data=trainData, method="rf", trControl=controlRf, ntree=250)
modelRf
```
Performance  and final analysis 
  
```{r, cache = T}
result <- predict(modelRf, testCleaned[, -length(names(testCleaned))])
result
```  


```{r, cache = T}
corrPlot <- cor(trainData[, -length(names(trainData))])
corrplot(corrPlot, method="color")
```

```{r, cache = T}
treeModel <- rpart(classe ~ ., data=trainData, method="class")
prp(treeModel) # fast plot
```